---
title: "Fijian Earthquakes"
author: "Jay Lee"
date: "9/7/2017"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
data(quakes)
library(ggplot2)
library(dplyr)
library(leaflet)
set.seed(365)
```

### Earthquake Detection

1. 

```{r}
ggplot(quakes, aes(x = mag, y = stations)) +
  geom_point() +
  geom_jitter()
```

The data shows a strong, positive, linear relationship.


2. If there was no relationship between these two values, we would expect the slope of the model to be 0 and the y-intercept to be $\bar{y}$.


3. 

```{r}
m1 <- lm(stations ~ mag, quakes)
summary(m1)
ggplot(quakes, aes(x = mag, y = stations)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = "lm", se = F)
```

For every increase of 1 in magnitude, we would expect a quake to be felt by 46.2822 more stations.

4. 

```{r echo = T}
x_bar <- mean(quakes$mag)
y_bar <- mean(quakes$stations)
beta_1 <- sum((quakes$mag - x_bar) * (quakes$stations - y_bar)) / sum((quakes$mag - x_bar)^2)
beta_0 <- mean(quakes$stations)-beta_1*mean(quakes$mag)
beta_1
beta_0
```


5. 

```{r echo = T}
lower <- 46.2822 - 1.962*.9034
upper <- 46.2822 + 1.962*.9034
lower
upper
confint(m1, "mag")
```

6. 

```{r}
46.2822*7.0-180.4243
```


7. Data description, inference, inference, inference, prediction, prediction

### Simulation

8. 

```{r}
set.seed(385)
sim_data <- sample_n(quakes, 1000, replace = T) %>%
  select(mag)
head(sim_data)
```


9. 

```{r}
f_hat <- function(x) {
  46.2822*x-180.4243
}
sim_data <- sim_data %>%
  mutate(y_hat = f_hat(mag))
head(sim_data)
```


10. 

```{r}
set.seed(1738)
RSS <- sum((m1$residuals)^2)
epsilon <- rnorm(1000, mean = 0, sd = sqrt(RSS/(1000-2)))
y <- sim_data$y_hat + epsilon
sim_data <- data.frame(sim_data, epsilon, y) %>%
  mutate(y = ifelse(y < 0, 0, y)) # can't be detected by a negative number of stations
head(sim_data)
```


11. 

```{r}

ggplot(sim_data, aes(x = mag, y = y)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = "lm", se = F) +
  geom_abline(slope = beta_1, intercept = beta_0, color = "red")
```

The simulated data set appears to match the original data set well. My simulated points (1000 uniform random draws from the original vector of `mag`) don't appear to have quite as many values for `mag` between 4.0 and 4.5 as the original, so there are fewer points there. This is reflected somewhat in the means of the two `mag` vectors.

```{r echo = T}
mean(quakes$mag)
mean(sim_data$mag)
```

If we form a linear model (below) of the simulated data, we see that both of the simulated model's parameters are inside the 95% confidence intervals for our original model. The original model is plotted in red above, and the simulated model in blue.

```{r echo = T}
sim <- lm(y~mag, sim_data)
coefficients(sim)
confint(m1, "mag")
confint(m1, "(Intercept)")
```


### Challenge Problem

```{r echo = T}
map <- leaflet(quakes) %>%
  addProviderTiles(providers$OpenMapSurfer.Roads) %>%
  addCircleMarkers(radius = ~mag*10-39, fill = F, weight = 1, opacity = ~stations/200+.1, color = "darkgreen", label = ~as.character(mag))
map
```

