---
title: "I trust I can rely on your vote"
author: "Jay Lee"
date: "11/19/2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
library(ggplot2)
library(voteogram)
poverty <- read.delim("https://raw.githubusercontent.com/andrewpbray/math-243/master/assets/data/poverty.txt", header = TRUE)
```

## Part I

### 1

```{r}
km2 <- kmeans(poverty[ ,-1], 2, nstart = 25)
kd2 <- data.frame(poverty$State, "k" = km2$cluster)

km3 <- kmeans(poverty[ ,-1], 3, nstart = 25)
kd3 <- data.frame(poverty$State, "k" = km3$cluster)

km4 <- kmeans(poverty[ ,-1], 4, nstart = 25)
kd4 <- data.frame(poverty$State, "k" = km4$cluster)

different_k <- left_join(kd2, kd3, by = "poverty.State") %>%
  left_join(kd4, by = "poverty.State") %>%
  arrange(poverty.State)
colnames(different_k) <- c("state", "k_2", "k_3", "k_4")

kable(different_k)
```

### 2

```{r}
pca_pov <- prcomp(poverty[, -1])
pov_pcs <- data.frame(PC1 = pca_pov$x[, 1],
                PC2 = pca_pov$x[, 2],
                cluster_2 = as.factor(km2$cluster),
                cluster_3 = as.factor(km3$cluster),
                cluster_4 = as.factor(km4$cluster),
                state = poverty$State)
ggplot(pov_pcs, aes(x = PC1, y = PC2)) +
  geom_point() +
  geom_text(aes(label = state), vjust = 2)
```

### 3

```{r}
ggplot(pov_pcs, aes(x = PC1, y = PC2, col = cluster_2)) +
  geom_point() +
  geom_text(aes(label = state), vjust = 2)
ggplot(pov_pcs, aes(x = PC1, y = PC2, col = cluster_3)) +
  geom_point() +
  geom_text(aes(label = state), vjust = 2)
ggplot(pov_pcs, aes(x = PC1, y = PC2, col = cluster_4)) +
  geom_point() +
  geom_text(aes(label = state), vjust = 2)
```

### 4

```{r}
km1 <- kmeans(poverty[ ,-1], 1, nstart = 25)
km5 <- kmeans(poverty[ ,-1], 5, nstart = 25)

scree <- data.frame("K" = 1:5,
                 "SS" = c(km1$tot.withinss,
                          km2$tot.withinss,
                          km3$tot.withinss,
                          km4$tot.withinss,
                          km5$tot.withinss))
ggplot(scree, aes(x = K, y = SS)) +
  geom_point() +
  geom_line()
```

There's no clear "elbow" in the scree plot, indicating that there's not really a best number of clusters.

## Part II

### 1

```{r}
cut_cut_cut <- roll_call(critter = "house", 115, session = 1L, rcall = 637)
cut_cut_cut$description
cut_cut_cut$date_of_vote
cut_cut_cut$result
OR <- filter(cut_cut_cut$votes, state_abbrev == "OR")
kable(OR[ , c(3, 5, 6, 8, 9)])
```

* The name of the bill was the Tax Cut and Jobs Act.
* It was voted upon on 11/16/2017.
* The bill passed the House.
* Oregon's 5 representatives voted 4-1 on party lines against the bill.

```{r warnings = FALSE, message = FALSE}
house_carto(cut_cut_cut, pp_square=TRUE) +
  labs(x=NULL, y=NULL, 
       title="House Vote 637 - Passes $1.5 Trillion Tax Cut") +
  hrbrthemes::theme_ipsum(plot_title_size = 24) +
  theme_voteogram()
```


### 2

```{r}
votes <- matrix(NA, nrow = 435, ncol = 50)
rcall_ind <- 588:637

for(i in 1:length(rcall_ind)) {
  votes[, i] <- roll_call("house", 115, 1, rcall_ind[i])$votes$position
}

vote_yes <- votes == "Yes"

dim(votes)
```

`votes` is 435 observations (representatives) by 50 variables (distinct bil votes). `vote_yes` is just `votes` recoded so that voting "Yes" is `TRUE` and any other action ("No", "Not Voting") is `FALSE`.

### 3

```{r}
pol2 <- kmeans(vote_yes, 2, nstart = 25)
table(pol2$cluster, cut_cut_cut$votes$party)
```

For the most part the clustering identifies party lines.

### 4

```{r}
pol1 <- kmeans(vote_yes[ ,-1], 1, nstart = 25)
pol3 <- kmeans(vote_yes[ ,-1], 3, nstart = 25)
pol4 <- kmeans(vote_yes[ ,-1], 4, nstart = 25)
pol5 <- kmeans(vote_yes[ ,-1], 5, nstart = 25)
pol6 <- kmeans(vote_yes[ ,-1], 6, nstart = 25)

pol_scree <- data.frame("K" = 1:6,
                 "SS" = c(pol1$tot.withinss,
                          pol2$tot.withinss,
                          pol3$tot.withinss,
                          pol4$tot.withinss,
                          pol5$tot.withinss,
                          pol6$tot.withinss))
ggplot(pol_scree, aes(x = K, y = SS)) +
  geom_point() +
  geom_line()
```

### 5

```{r}
pca_pol <- prcomp(vote_yes)
pol_pcs <- data.frame(rep = cut_cut_cut$votes$member_name,
                      PC1 = pca_pol$x[, 1],
                      PC2 = pca_pol$x[, 2],
                      cluster_4 = as.factor(pol4$cluster),
                      party = cut_cut_cut$votes$party)
ggplot(pol_pcs, aes(x = PC1, y = PC2, color = cluster_4)) +
  geom_point() +
  geom_text(aes(label = party), vjust = 2)
```

### 6

The first several loadings can tell us in which directions the data has the most variance, i.e. in which directions the data points are most different from each other.

### 7

```{r}
table(pol_pcs$party, pol_pcs$PC1 < 0)
kable(head(arrange(pol_pcs, PC1, PC2), 50))
```

Together, the first two components do a good job of separating Repubicans from Democrats. I can't really tell what they do individually.